---
title: "data_wrangling_1_9-20-18"
author: "J Shearston"
date: "September 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import my first couple of csv files.

```{r}
litters_data = read_csv(file = "./data_import_examples/data_import_examples/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
names(litters_data)

pups_data = read_csv(file = "./data_import_examples/data_import_examples/FAS_pups.csv", col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
names(pups_data)

```

## Look at Imported Data

Look throuh imported datasets.

```{r}
head(litters_data, 5)
tail(litters_data, 5)
skimr::skim(litters_data)

head(pups_data, 5)
tail(pups_data, 5)
skimr::skim(litters_data)

```

## Reading in other file formats

Read in MLB data (excel format)

```{r}
library(readxl)
mlb_data = readxl::read_excel(path = "./data_import_examples/data_import_examples/mlb11.xlsx")
mlb_subset = readxl::read_excel(path= "./data_import_examples/data_import_examples/mlb11.xlsx", 
                                range = "A1:E17")

```

Read in Pulse data (SAS format)

```{r}
library(haven)
pulse_data = read_sas("./data_import_examples/data_import_examples/public_pulse_data.sas7bdat")
head(pulse_data, 5)

```

##Select variables

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_survive)

```

...by specifying a range:

```{r}
select(litters_data, group, gd_of_birth:pups_survive)
```

...by removing what you want to remove

```{r}
select(litters_data, -litter_number)
```

...also rename as you select, or just rename

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, orig_litter_number = litter_number)
```

...pull in stuff that starts with something in particular, or reorder (select helpers)

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, everything())
```


##Filtering observations

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, group == "Low7" | group == "Low8")
filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)
```


##Mutate

Create variables

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
       )

mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2
       )
```


##Arrange

Putting data in a specific order, for example arranging by a specific variable numerically

```{r}
arrange(litters_data, gd0_weight)

arrange(litters_data, desc(pups_born_alive), gd0_weight)
```


##Piping

Look at intermediate object approach:

```{r}
litters_data_raw = read_csv("./data_import_examples/data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Use nested function calls

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data_import_examples/data_import_examples/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use piping

```{r}
read_csv("./data_import_examples/data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight))
  
```

Illustrate placeholder use with `lm`

```{r}
read_csv("./data_import_examples/data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight)) %>%
  lm(gd18_weight ~ gd0_weight, data = .)

```

